---
title: "Corporate Finance Assignment 4"
author: "Nitish Ramkumar"
output: html_document
---
```{r echo=FALSE}
# Function to get data #
getData <- function(sql, n = -1){
  #setup connection
  res <- dbSendQuery(wrds, sql)
  dbHasCompleted(res)
  
  #perform fetch
  returnData <- fetch(res, n)
  
  #clear memory
  dbClearResult(res)
  return(returnData)
}
```
  
For the annual dataset on earnings, cashflows and leverage ratios, we should take the necessary fundamentals from **Compustat**. The file **COMPM.FUNDA** contains all the necessary information. The earnings and cashflow data from Compustat are in the units of millions.  
  
The following formulas were used for calculation of the required information.  
  
*Total Cashflow = Operating Cashflow + Investing Cashflow + Financing Cashflow*  
  
*Market Leverage Ratio = (Current Debt + Long term debt)/Total Assets*  
  
*Book Leverage Ratio = (Current Debt + Long term Debt)/(Current Debt + Long term debt + Common Equity)*  
  
*Earnings = Net Income*  
  
**Sample of the resultant data**
```{r echo=FALSE}
#data from WRDS
sql.fundamentals <- "SELECT GVKEY AS GVKEY, DATADATE AS Date,OANCF AS OperatingCF,
                            IVNCF AS InvestingCF,FINCF AS FinancingCF,NI as Earnings,
                            DLC AS CurrentDebt, DLTT AS LongTermDebt, 
                            AT AS TotalAssets, SEQ AS CommonEq, PSTK AS LiquidationVal 
                            FROM COMPM.FUNDA funda"
fundamentals <- getData(sql.fundamentals)

sql.permnoLink <- "SELECT GVKEY,LPERMNO FROM CRSPA.CCMXPF_LINKTABLE"
permnoLink <- getData(sql.permnoLink)

suppressMessages(library(dplyr))

#Use dplyr to format data and join with link table to get permno
fundamental.formatted <- fundamentals %>% filter(complete.cases(.)) %>% mutate(CFInMil = OperatingCF + InvestingCF + FinancingCF, marketLRatio = (CurrentDebt+LongTermDebt)/TotalAssets, bookLRatio = (CurrentDebt+LongTermDebt)/(CurrentDebt+LongTermDebt+CommonEq))%>%select(GVKEY,Date,CFInMil,EarningsInMil = Earnings,marketLRatio,bookLRatio)

fundamental.formatted %>% left_join(permnoLink %>% na.omit(),by= c("GVKEY"="gvkey")) %>% head
```
  
  
**R Code**
```{r eval=FALSE}
sql.fundamentals <- "SELECT GVKEY AS GVKEY, DATADATE AS Date,OANCF AS OperatingCF,
                            IVNCF AS InvestingCF,FINCF AS FinancingCF,NI as Earnings,
                            DLC AS CurrentDebt, DLTT AS LongTermDebt, 
                            AT AS TotalAssets, SEQ AS CommonEq, PSTK AS LiquidationVal 
                            FROM COMPM.FUNDA funda"
fundamentals <- getData(sql.fundamentals)

sql.permnoLink <- "SELECT GVKEY,LPERMNO FROM CRSPA.CCMXPF_LINKTABLE"
permnoLink <- getData(sql.permnoLink)

suppressWarnings(library(dplyr))

fundamental.formatted <- fundamentals %>% filter(complete.cases(.)) %>% mutate(cashflowInMil = OperatingCF + InvestingCF + FinancingCF, marketLRatio = (CurrentDebt+LongTermDebt)/TotalAssets, bookLRatio = (CurrentDebt+LongTermDebt)/(CurrentDebt+LongTermDebt+CommonEq))%>%select(GVKEY,Date,cashflowInMil,EarningsInMil = Earnings,marketLRatio,bookLRatio)

fundamental.formatted %>% left_join(permnoLink %>% na.omit(),by= c("GVKEY"="gvkey")) %>% head

```