---
title: "Corporate Finance Assignment 5"
author: "Nitish Ramkumar"
output: html_document
---
#Assignment 3

###Dataset building

To compare, lets build a PERM,YYYYMMDD,Ret table which has PERMNO starting with only 100. Sample of the merged dataset is:
```{r echo=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(lubridate))
load("C:/_UCLA/237A_CorporateFinance/Assignments/MFE_237A_CorpFinance/Assignment2/total_data_withForrest.RData")

dataSetForComparision2_1 <- stock.annualReturns.WithTicker.Final %>% filter( PERMNO %/% 100 == 100 & date>='2013-01-01' & date<'2015-01-01')%>%select(PERMNO,date,AnnualReturn,MarketCap)
dataSetForComparision2_1$date <- as.numeric(format(dataSetForComparision2_1$date,"%Y%m%d"))

raw[,"Year"] <- paste0(raw[,"Year"],"12","31")
raw[,"PERMNO"] <- as.numeric(raw[,"PERMNO"])
raw[,"Year"] <- as.numeric(raw[,"Year"])
raw[,"Return"] <- as.numeric(raw[,"Return"])
raw[,"Market Cap"] <- as.numeric(raw[,"Market Cap"])
dataSetForComparision2_2 <- raw[,c("PERMNO","Year","Return","Market Cap")]
dataSetForComparision2_2 <- apply(dataSetForComparision2_2,2,unlist)
dataSetForComparision2_2 <- dataSetForComparision2_2[dataSetForComparision2_2[,"PERMNO"]%/%100==100,]
dataSetForComparision2_2 <- data.frame(dataSetForComparision2_2,stringsAsFactors = FALSE,row.names = NULL)
colnames(dataSetForComparision2_2) <- c("PERMNO","date","AnnualReturn","MarketCap")

joined <- dataSetForComparision2_1 %>% inner_join(dataSetForComparision2_2,by = c("PERMNO","date"))
head(joined)
```
x is my data, and y is Forrest's data.  

###Comparison Analysis
**Returns**  
  
If we see the plot of returns between the annualized returns of me and Forrest, we notice that it is not along the 45 degree line (green line).    
```{r echo=FALSE}
plot(joined$AnnualReturn.x~joined$AnnualReturn.y,xlab="Forrest Val",ylab = "My val",main="Annualized returns comparison")
abline(0,1,col="green")
```

This difference in values is because Forrest has correctly maintained the log of returns in the Annual return column, whereas I have maintained the actual returns.

**MarketCap**  
  
I didn't multiply the value by 1000, which Forrest has correctly done. If we plot the market cap between the two datasets,  

```{r echo=FALSE}
plot(joined$MarketCap.x*1000~joined$MarketCap.y,xlab="Forrest Val",ylab = "My val",xlim=c(30000000,1598000000),main="Annualized Marketcap value")
abline(0,1,col="green")
```

One reason I can point out for this is that I have taken annual market cap as average of all monthly market caps. Forrest has maintained the year end value as marketcap.  
  
###Reconciliation 
1) Both of us should use log returns
2) Use same standard for market cap value (either average or end of year value)
